
# Walking preference {#sec-preference}
While quantitative walkability indicators measure the "potential for walking", data sources like travel diaries collect real-world preferences. Walking trips from three years of the New Zealand Household Travel Survey (HTS) were routed on the Wellington network to show preferential areas of the city for pedestrian movement. 


```{r}
library(sf)
library(dplyr)
library(janitor)
library(haven)
library(purrr)
library(tidyr)
library(tmap)
library(kableExtra)
```

```{r}
#| output: false
wgtn_lines <- st_read("../../wgtn_network/wgtn_lines.shp") |> 
  dplyr::select(-slp_cls)

wgtn_lines$slope_class = wgtn_lines$slope %>%
  cut(
    breaks = c(0, 3, 5, 8, 10, 20, Inf),
    labels = c("0-3: flat", "3-5: mild", "5-8: medium", "8-10: hard", 
               "10-20: extreme", ">20: impossible"),
    right = F
  )
```

```{r}
#| output: false
routes <- st_read("../../routes_3years_data/routes.shp") |> 
  mutate(slp_cls = factor(
    slp_cls, 
    levels = c(
      "0-3: flat", 
      "3-5: mild", 
      "5-8: medium", 
      "8-10: hard", 
      "10-20: extreme",
      ">20: impossible"
      )
    )
    ) |> 
  arrange(slp_cls) 
```

Streets with walking trips are strongly biased (70%) towards flat or mild gradients. The prevalence of these low gradient streets is only 44% in the whole of the city (@tbl-all-streets) with even lower prevalence of 33% for residential streets (@tbl-res-streets). 

```{r}
# summary - all
all_summary <- routes |> 
  st_set_geometry(NULL) |> 
  filter(!is.na(slp_cls)) |> 
  group_by(slp_cls) |> 
  summarise(n = sum(length)) |> 
  mutate(percentage = round((n / sum(n)) * 100, 1)) |> 
  dplyr::select(-n)

all_summary |> 
  rename(`%` = percentage,
         `Slope` = slp_cls) |> 
  kable()
```

